Convex Surrogates via Polyhedral Losses

### 2019-01-30 cvx-fin.tex
Cleaning up appendix

\begin{definition}\label{def:trim}
  Let $\Gamma:\simplex \toto\R'$ be an elicitable property.
  We define $\trim(\Gamma)$ as the set of maximal \botodo{suggest ``maximal'' instead of ``full-dimensional''-- changed} level sets of $\Gamma$.
  Formally, $\trim(\Gamma) = \{\Gamma_u : u \in \R'\} \setminus \{ \Gamma_u : \exists w \in \R' \text{ s.t. } \Gamma_u \subsetneq \Gamma_w \}$.
\end{definition}
\raf{Explain why this is a useful concept in general, and for convex elic}  
Looking at the $\trim$ of a property allows us to relate the finite properties we are interested in and the subtlely different surrogate properties that we actually elicit when we embed the original property into $\reals^d$.
In the surrogate property, there is often an infinite set of ``hidden'' level sets corresponding to the reports in the convex hull of embedded points.
Taking the $\trim$ thus allows us to remove labels (thus also removing redundant reports) so that we can relate the optimal sets of original reports and their embedded points, and then we remove these ``hidden'' level sets that are only optimal when we are indifferent between a subset of the embedded points.

\begin{proposition}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}
  \item $\Gamma$ embeds a finite property $\gamma:\simplex \toto \R$.
    \raf{I think we don't need elicitability of $\gamma$ for the TFAE, right?  If so, let's remove it here (just ``finite property'') and add it to the moreover (``, and $\gamma$ is elicitable.'')  The proof that $\trim(\Gamma)$ (with some labels) is elicitable I think is in my notes.}\jessie{That sounds good to me, as long as $\Gamma$ is still elicitable.  Should ``nonredundant'' be added to the moreover as well?}
  \item $\trim(\Gamma)$ is a finite set.
  \item There is a finite set of maximal level sets $\Theta$ of $\Gamma$ that union to $\simplex$.
  \end{enumerate}
  Moreover, when any of the above hold, $\{\gamma_r : r\in\R\} = \trim(\Gamma) = \Theta$, and $\gamma$ is elicitable.
\end{proposition}
\begin{proof}
  For $1 \implies 2$, since $\Gamma$ embeds the finite property $\gamma$, we show that $\trim(\Gamma) = \{\Gamma_u : u \in \varphi(\R) \} = \{\gamma_r : r \in \R\}$.
  (Note that the second equality follows from the definition of embedding.)
  Clearly, $\trim(\Gamma) \supseteq \{\Gamma_u : u \in \varphi(\R)\}$ as we are simply restricting the number of level sets we are indexing over.
  Now, consider some $\theta \in \trim(\Gamma)$ and the report $u$ so that $\theta = \Gamma_u$.
  If $u \in \varphi(\R)$, then clearly we are done.
  If $u \not\in \varphi(\R)$ but there is some report $w \in \varphi(\R)$ so that $\Gamma_u = \Gamma_w$, then we are again done since $\trim(\Gamma)$ is nonredundant.
  If there is no such $w \in \R$, then we claim $\Gamma_u \not \in \trim(\Gamma)$.
  To see this, consider that if $\Gamma_u$ is full-dimensional, it must be a relabeling of another report by Lemma~\ref{lem:unique-opt-on-inter} in the Appendix.
  If $\Gamma_u$ is not maximal, it must be a proper subset of some maximal level set by Lemma~\ref{lem:trim-subsets-not-maximal}, and therefore not in $\trim(\Gamma)$.
%\jessiet{skimmed over this, but follows from the convexity of the scoring rule.  My thought for the paper is to put that proof in the appendix,and just have a footnote or side note here with a reference to the proof in the appendix.}
  Thus, we conclude $\trim(\Gamma) \subseteq \{\Gamma_u : u \in \varphi(\R)\}$ and since the sets are equal, can conclude that if $\Gamma$ embeds the finite property $\gamma$, then $\trim(\Gamma)$ is finite.

  For $2 \implies 3$, since each level set $\theta \in \trim(\Gamma)$ is full-dimensional by Lemma~\ref{lem:trim-full-dim}, suppose for now that $\Theta := \trim(\Gamma)$.
  %As $\Gamma$ is nondegenerate, for every $p \in \simplex$, there must be some report $u$ corresponding to the level set such that $p \in \Gamma_u$.
  Since $\Gamma$ is nondegenerate, for all $p \in \simplex$, there is a level set $\theta \in \trim(\Gamma)$ such that $p \in \theta$, since the level sets of elicitable properties are closed and convex, and only non-full-dimensional level sets are removed in taking $\trim(\Gamma)$.
  If, for some $p\in \simplex$, the level set $\Gamma_u$ corresponding to $p$ was removed, it would be because there is a level set corresponding to $w \in \reals^d$ such that $\Gamma_u \subset \Gamma_w$, but then $p \in \Gamma_w$.

  For $3 \implies 1$, take the set $\R := \{u : \Gamma_u \in\trim(\Gamma)\}$.
  (Take only one such report per level set.)
  The set $\R$ is then finite, and taking $\varphi$ to be the identity, we observe that $\Gamma$ embeds $\gamma: p \mapsto \Gamma(p) \cap \R$.
  %We can also see $\gamma = \prop{L|_{\R}}$, where $\Gamma = \prop{L}$.
\end{proof}

@@

\section{Considering unlabeled properties}
  Note that $\Gamma$ is elicitable, which means that there is a convex score $G:\simplex \to \reals$ representing the Bayes Risk (i.e. $G(p)$ is the minimal expected loss over all reports at $p$.)
  There is a set $\D \subseteq \partial_p G$ so that there is a bijection $\phi:\D \to \R$.
  Therefore, the subgradients of $G$ correspond to the property value at the distribution $p$.

@@

\begin{lemma}\label{lem:trim-subsets-not-maximal}
Let $\Gamma:\simplex \toto \R'$ be an elicitable property.
For any $u,w \in \R'$, if $\Gamma_u \subsetneq \Gamma_w$, then $\Gamma_u$ is not full dimensional.
\end{lemma}
\begin{proof}
  If there is a maximal level set $\Gamma_u$ that is a proper subset of $\Gamma_w$, then the corresponding scoring rule $G$ must not be differentiable (i.e. $|\partial_p G(p)| > 1$) for every distribution $p \in \Gamma_u$.
  However, closed, convex, maximal level sets have positive Lebesgue measure in $\reals^n$, and thus $G$ can not be differentiable on $\Gamma_u$.
  This contradicts the convexity of $G$, as convex functions are differentiable almost everywhere, which in turn contradicts the elicitability of $\Gamma$.
  Therefore, we conclude that $\Gamma_u$ is not maximal.
\end{proof}

\begin{lemma}\label{lem:unique-opt-on-inter}
  Let $\Gamma:\simplex \toto \R'$ be a nonredundant, elicitable property.
  For all $u \in \R'$, if $p \in \inter{\Gamma_u}$, then $\Gamma(p) = \{u\}$.
\end{lemma}
\begin{proof}
\jessiet{Come back and formalize}
  We'll use the relationship between nonredundant properties and subgradients of the expected score $G$.
  Now consider the following convex analysis fact: for any convex function and any two point which share a subgradient, the function is flat between those points.
  In particular, $G$ is flat on $\Gamma_u$, and therefore $G$ must be differentiable on the interior of $\Gamma_u$.
  Therefore, $G$ must only have one subgradient on $\inter{\Gamma_u}$.
\end{proof}

\begin{corollary}\label{cor:cant-span-interiors}
  If $\Gamma:\simplex \toto \R'$ is a nonredundant, finite elicitable property, no level set spans the interior of two other level sets.
  Formally, there is no $w \in \R'$ so that for any $p \in \inter{\Gamma_u}$ and $q \in \inter{\Gamma_v}$, $p,q \in \Gamma_w$ for reports $u,v \in \R'$.
\end{corollary}

\begin{lemma}\label{lem:trim-full-dim}
  Consider a nondegenerate, elicitable property $\Gamma$ with finitely many level sets.
  Every $\theta \in \Theta := \trim(\Gamma)$ is full-dimensional.  
\end{lemma}
\begin{proof}
  Consider an arbitrary level set $\theta$.
  If $\theta$ is not maximal, we want to show that it must be the intersection of some maximal level sets in $\trim(\Gamma)$.\jessiet{Why?  Add more details about why the level set can't span two maximal level sets}
  If it is a proper subset of the level sets of which it is the intersection, and so it is not in the $\trim$ as it is in the subtracted set.

  Now suppose that a level set $\theta$ was not full-dimensional and not a proper subset of any one level set.
  It must then span multiple level sets, which, by Corollary~\ref{cor:cant-span-interiors} contradicts the elicitability of $\Gamma$.
  Thus, any $\theta$ that is not full-dimensional must be the intersection of a finite set of full-dimensional level sets $\theta_i \in \trim(\Gamma)$, and therefore $\theta \not \in \trim(\Gamma)$.
  Therefore, each $\theta \in \trim(\Gamma)$ must be full-dimensional.
\end{proof}

>>>>

\begin{definition}\label{def:trim}
  Let $\Gamma:\simplex \toto\R'$ be an elicitable property.
  We define $\trim(\Gamma)$ as the set of maximal \botodo{suggest ``maximal'' instead of ``full-dimensional''-- changed} level sets of $\Gamma$.
  Formally, $\trim(\Gamma) = \{\Gamma_u : u \in \R'\} \setminus \{ \Gamma_u : \exists w \in \R' \text{ s.t. } \Gamma_u \subsetneq \Gamma_w \}$.
\end{definition}
  Take note that the unlabeled property $\trim(\Gamma)$ is nonredundant, meaning that for any $\theta \in \trim(\Gamma)$, there is no level set $\theta' \in \trim(\Gamma)$ such that $\theta \subset \theta'$.

\raf{Explain why this is a useful concept in general, and for convex elic}  
  Looking at the $\trim$ of a property allows us to relate the finite property of interest to the subtlely different surrogate property that we actually elicit when embedding the original property into $\reals^d$.
  In the surrogate property, there is often an infinite set of ``hidden'' level sets corresponding to the reports in the convex hull of embedded points whose level sets have nonempty intersection.
  Taking the $\trim$ thus allows us to remove labels so that we can relate the optimal sets of original reports and their embedded points, and then we remove the ``hidden'' level sets that are only optimal when we are indifferent between a subset of the embedded points.

\begin{proposition}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}
  \item $\Gamma$ embeds a finite property $\gamma:\simplex \toto \R$.
    \raf{I think we don't need elicitability of $\gamma$ for the TFAE, right?  If so, let's remove it here (just ``finite property'') and add it to the moreover (``, and $\gamma$ is elicitable.'')  The proof that $\trim(\Gamma)$ (with some labels) is elicitable I think is in my notes.}\jessie{That sounds good to me, as long as $\Gamma$ is still elicitable.  Should ``nonredundant'' be added to the moreover as well?-- No, because you could have a redundant, finite property so that these are true.}
  \item $\trim(\Gamma)$ is a finite set.
  \item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$ that union to $\simplex$.
  \end{enumerate}
  Moreover, when any of the above hold, $\{\gamma_r : r\in\R\} = \trim(\Gamma) = \Theta$, and $\gamma$ is elicitable.
\end{proposition}
\begin{proof}
  For $1 \implies 2$, since $\Gamma$ embeds the finite property $\gamma$, we show that $\trim(\Gamma) = \{\Gamma_u : u \in \varphi(\R) \} = \{\gamma_r : r \in \R\}$.
  (Note that the second equality follows from the definition of embedding.)
  Clearly, $\trim(\Gamma) \supseteq \{\Gamma_u : u \in \varphi(\R)\}$ as we are simply restricting the number of level sets we are indexing over.
  Take some $\theta \in \trim(\Gamma)$ and consider the report $u$ so that $\theta = \Gamma_u$.
  If $u \in \varphi(\R)$, then clearly the subset inclusion is true.
  If $u \not\in \varphi(\R)$ but there is some report $w \in \varphi(\R)$ so that $\Gamma_u = \Gamma_w$, then we are again done.

  If there is no such $w \in \R$, then we claim $\Gamma_u \not \in \trim(\Gamma)$.
  To see this, consider that if $\Gamma_u$ is full-dimensional, it must be a relabeling of another report by Lemma~\ref{lem:unique-opt-on-inter} in the Appendix.
  If $\Gamma_u$ is not full-dimensional, it must be a proper subset of some full-dimensional level set by Lemma~\ref{lem:trim-subsets-not-full-dimensional}, and therefore not in $\trim(\Gamma)$.
%\jessiet{skimmed over this, but follows from the convexity of the scoring rule.  My thought for the paper is to put that proof in the appendix,and just have a footnote or side note here with a reference to the proof in the appendix.}
  Thus, we conclude $\trim(\Gamma) \subseteq \{\Gamma_u : u \in \varphi(\R)\}$.
  Since the sets are equal and the latter is finite, we conclude that if $\Gamma$ embeds the finite property $\gamma$, then $\trim(\Gamma)$ is finite.

  For $2 \implies 3$, since each level set $\theta \in \trim(\Gamma)$ is full-dimensional by Lemma~\ref{lem:trim-full-dim}, suppose for now that $\Theta := \trim(\Gamma)$.
  %As $\Gamma$ is nondegenerate, for every $p \in \simplex$, there must be some report $u$ corresponding to the level set such that $p \in \Gamma_u$.
  Since $\Gamma$ is nondegenerate, for all $p \in \simplex$, there is a level set $\theta \in \trim(\Gamma)$ such that $p \in \theta$, since the level sets of elicitable properties are closed and convex, and only level sets that are not full-dimensional level sets are removed in taking $\trim(\Gamma)$.
  If, for some $p \in \simplex$, the level set $\Gamma_u$ corresponding to $p$ was removed, it would be because there is a level set corresponding to $w \in \reals^d$ such that $\Gamma_u \subset \Gamma_w$, but then $p \in \Gamma_w$.

  For $3 \implies 1$, take the set $\R := \{u : \Gamma_u \in \Theta\}$.
  (Take only one such report per level set.)
  The set $\R$ is then finite, and taking $\varphi$ to be the identity, we observe that $\Gamma$ embeds $\gamma: p \mapsto \Gamma(p) \cap \R$.
  %We can also see $\gamma = \prop{L|_{\R}}$, where $\Gamma = \prop{L}$.
\end{proof}

@@

\section{Considering unlabeled properties}
  Note that when $\Gamma$ is elicitable, there is a convex score $G:\simplex \to \reals$ representing the Bayes Risk (i.e. $G(p)$ is the minimal expected loss over all reports at $p$.)
  There is a set $\D \subseteq \partial_p G$ so that there is a bijection $\phi:\D \to \R$.
  Therefore, the subgradients of $G$ correspond to the property value at the distribution $p$.

@@

\begin{lemma}\label{lem:trim-subsets-not-full-dimensional}
Let $\Gamma:\simplex \toto \R'$ be an elicitable property.
For any $u,w \in \R'$, if $\Gamma_u \subsetneq \Gamma_w$, then $\Gamma_u$ is not full dimensional.
\end{lemma}
\begin{proof}
  If there is a full-dimensional level set $\Gamma_u \subsetneq \Gamma_w$, then the corresponding scoring rule $G$ must not be differentiable (i.e. $|\partial_p G(p)| > 1$) for all $p \in \Gamma_u$.
%  (This is true because $\phi^{-1}(u) \neq \phi^{-1}(w)$ as the level sets are not equal.)
  However, closed, convex, full-dimensional level sets have positive Lebesgue measure in $\reals^{n-1}$,\jessiet{I don't know if we need a citation/proof for this assertion...} and thus contradicts the convexity of $G$, as convex functions are differentiable at all but a countable set of points.
  This in turn contradicts the elicitability of $\Gamma$, and so, we conclude that $\Gamma_u$ is not full-dimensional.
\end{proof}

\begin{lemma}\label{lem:unique-opt-on-inter}
  Let $\Gamma:\simplex \toto \R'$ be an elicitable property such that $\trim(\Gamma)$ is finite.
  For all $u \in \R'$, if $p \in \inter{\Gamma_u}$, then $\Gamma(p) = \{u\}$.
\end{lemma}
\begin{proof}
\jessiet{Come back and formalize}
  We'll use the relationship between nonredundant properties and subgradients of the expected score $G$.
  Now consider the following convex analysis fact: for any convex function and any two point which share a subgradient, the function is flat between those points.
  In particular, $G$ is flat on $\Gamma_u$, and therefore $G$ must be differentiable on the interior of $\Gamma_u$.
  Therefore, $G$ must only have one subgradient on $\inter{\Gamma_u}$.
  \bigskip
  \jessie{Formalization below}
  \hline
  Consider that $\trim(\Gamma)$ is nonredundant, and $p \in \inter{\Gamma_u}$ only if $\Gamma_u \in \trim(\Gamma)$ by Lemma~\ref{lem:trim-full-dim}, as a closed, convex set is full-dimensional if and only if it has nonempty interior. \jessiet{Sanity check here?}
  Since $\Gamma$ is elicitable, the score $G$ must be convex.
  For $w \in \R'$, if $\Gamma_w \subseteq \Gamma_u$, then we contradict the nonredundancy of $\trim(\Gamma)$, so we only worry about the case where $\Gamma_w \not \subseteq \Gamma_u$..

  Suppose there is such a level set.
  There must be another level set $v$ so that $\Gamma_w \cap \inter{\Gamma_v}$ is nonempty, since level sets of elicitable properties are closed.
  Therefore, for $q \in \Gamma_w$, we observe $\Gamma(q) \subseteq \{u,w\}$ or $\{v,w\}$.
  More importantly- for $q \in \Gamma_w$, $\Gamma(q)$ is not constant.

  Let's say there is a $q \in \Gamma_w \cap \inter{\Gamma_u}$ and $\q' \in \Gamma_w \cap \inter{\Gamma_v}$.
  Additionally, $\conv(\{q,q'\}) \subseteq \Gamma_w$, and observe that $\conv(\{q,q'\})$ is not countable.
  Since there is a bijection from $\Gamma(p)$ to $\partial_p G(p)$, we then note that $|\partial_pG(p)| > 1$ on an uncountable set, and $\partial_p G(p)$ is not constant on $\conv(\{q,q'\})$.
  This contradicts the convexity of $G$, and in turn, the elicitability of $\Gamma$ and $\trim(\Gamma)$.

  Therefore, for all $p \in \inter{\Gamma_u}$, we must observe $\Gamma(p) = \{u\}$.

\begin{corollary}\label{cor:cant-span-interiors}
  If $\Gamma:\simplex \toto \R'$ is a nonredundant, finite elicitable property, no level set spans the interior of two other level sets.
  Formally, there is no $w \in \R'$ so that for any $p \in \inter{\Gamma_u}$ and $q \in \inter{\Gamma_v}$, $p,q \in \Gamma_w$ for reports $u,v \in \R'$ with $u \neq v$.
\end{corollary}
\jessiet{I can write something for a proof of this if you want, but it follows from the previous statement pretty quickly.}

\begin{lemma}\label{lem:trim-full-dim}
  Consider a nondegenerate, elicitable property $\Gamma$ with finitely many full-dimensional level sets.
  Every $\theta \in \Theta := \trim(\Gamma)$ is full-dimensional.  
\end{lemma}
\begin{proof}
  Consider an arbitrary level set $\theta$.
  If $\theta$ is not full-dimensional, we want to show that it must be the intersection of some full-dimensional level sets in $\trim(\Gamma)$.\jessiet{Why?  Add more details about why the level set can't span two full-dimensional level sets}
  If it is a proper subset of the level sets of which it is the intersection, and so it is not in the $\trim$ as it is in the subtracted set.

  Now suppose that a level set $\theta$ was not full-dimensional and not a proper subset of any one level set.
  It must then span the interior multiple level sets, as level sets of elicitable properties are closed.
  By Corollary~\ref{cor:cant-span-interiors}, we contradict the elicitability of $\Gamma$.
  Thus, any $\theta$ that is not full-dimensional must be the intersection of a finite set of full-dimensional level sets $\theta_i \in \trim(\Gamma)$.
  As trimming a property removes such redundant reports, we then observe that $\theta \not \in \trim(\Gamma)$.
  Therefore, each $\theta \in \trim(\Gamma)$ must be full-dimensional.
\end{proof}
